### 2015-04-08 manuscript/correspondence/cover_letter.tex
Include Bioinformatics cover letter.

\documentclass[stdletter,letterpaper,addrfromleft,orderfromdateto,dateleft,11pt,noaddrto,sigleft]{newlfm}
\topmarginskip{-0.4in}
\bottommarginskip{-1.5in}
\leftmarginsize{1in}
\rightmarginsize{1.25in}
\sigskipbefore{0.2in}
\sigskipafter{0in}
\noLines
\nolines
\noHeadline
\noheadline
\signature{Trevor Bedford and Richard Neher}

\usepackage{graphicx}
\usepackage{hyperref}

\namefrom{}
\addrfrom{
	\includegraphics[width=6.5cm]{figures/fhcrc_logo} \\ 
	Vaccine and Infectious Disease Division \\ Fred Hutchinson Cancer Research Center \\ Seattle, WA, USA \\
	\\
	\includegraphics[width=4cm]{figures/max_planck_logo} \\ 
	Max Planck Institute for Developmental Biology \\ T\"ubingen, Germany}

\greetto{Dear Editor,}
\closeline{Sincerely,}

% comments
\usepackage{color} 
\usepackage{ulem}
\definecolor{purple}{rgb}{0.459,0.109,0.538}
\def\tb#1#2{\sout{#1} \textcolor{purple}{#2}} 
\def\tbc#1{\textcolor{purple}{[#1]}}

 \newcommand{\nextflu}{\textbf{nextflu}}

\begin{document}

\begin{newlfm}

Please find, attached, our application note entitled ``nextflu: Real-time tracking of seasonal influenza virus evolution in humans,'' which we would be grateful for your consideration for publication in \textit{Bioinformatics}.
We believe that \textit{Bioinformatics} is an appropriate venue to showcase our software application, as \nextflu{} represents a valuable contribution to public health bioinformatics as it allows better understanding of circulating viruses in an extremely timely fashion.
Source code to \nextflu{} is freely available from \url{https://github.com/blab/nextflu}, while a live web application is publicly available at \url{http://nextflu.org}.
At this point, we've received positive encouragement that \nextflu{} will be useful to several National Influenza Centers and WHO Collaborating Centers in aiding surveillance and strain selection efforts.
As such, we believe that \nextflu{} will see substantial adoption by the influenza research community.

\end{newlfm}
\end{document}